fGFPCA dynamic prediction

### 2023-01-26 f-GFPCA.Rmd
more notes
try but failed to get prediction interval

```{r}
source(here("Code/OutSampMLE.R"))
```

```{r}
# clean out-of-sample prediction results
out_pred_clean <- function(subjID, mid=mid){
  # Prediction
  pred_t195 <- out_samp_dyn_pred(df_new = df %>% filter(id==subjID & sind_inx<=195) %>%
                                  select(-eta_i),
                                fpca_fit = fpca_fit)
  pred_t395 <- out_samp_dyn_pred(df_new = df %>% filter(id==subjID & sind_inx<=395) %>%
                                  select(-eta_i),
                                fpca_fit = fpca_fit)
  pred_t595 <- out_samp_dyn_pred(df_new = df %>% filter(id==subjID & sind_inx<=595) %>%
                                  select(-eta_i),
                                fpca_fit = fpca_fit)

@@

```{r, fig.width=8, fig.height=4}
# compare point prediction with in-sample
# subject 1
grid.arrange(
  out_pred_subj1 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 1 out-of-sample"),

@@

# subject 2
grid.arrange(
  out_pred_subj2 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 2 out-of-sample"),

@@

# subject 3
grid.arrange(
  out_pred_subj3 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 3 out-of-sample"),

@@

# subject 4
grid.arrange(
  out_pred_subj4 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 4 out-of-sample"),

@@

@@

```{r}
# true scores:xi

# scores
in_score <- matrix(NA, nrow=N, ncol=4)
out_score <- matrix(NA, nrow=N, ncol=4)
for(i in 1:N){
  in_score[i, ] <- in_samp_dyn_pred(fpca_fit=fpca_fit, 
                   df_part=df_pred_unique %>% filter(id==i & bin<=595))$score
  out_score[i,] <- out_samp_dyn_pred(df %>% filter(id==i & sind_inx<=595) %>%
                                  select(-eta_i),
                                  fpca_fit=fpca_fit)$score_out
}

# bias
in_bias <- in_score-xi
out_bias <- out_score-xi
```

```{r, fig.height=4, fig.width=8}
grid.arrange(
  out_bias %>% data.frame() %>%
    pivot_longer(1:4, names_to = "PC", values_to = "Bias") %>%
    ggplot()+
    geom_boxplot(aes(x=PC, y=Bias))+
    stat_summary(aes(x=PC, y=Bias), fun = mean, geom = "point", col = "red")+
    labs(title = "Out-of-sample"),

@@

$$Var(\hat{\boldsymbol{\eta}}-\boldsymbol{\eta})=Var(\boldsymbol{\hat{f_0}-f_0})+\boldsymbol{\Phi\Gamma \Phi^T}+\hat{\sigma_{\epsilon}}^2$$
Note:

>>>>

```{r, results='hide'}
source(here("Code/OutSampMLE.R"))
source(here("Code/OutSampBayes.R"))
```

```{r, MLE_pred}
# clean MLE out-of-sample prediction results
out_pred_clean <- function(subjID, mid=mid){
  # Prediction
  pred_t195 <- out_samp_dyn_pred(df_new = df %>% filter(id==subjID & sind_inx<=195) %>%
                                  select(-eta_i),
                                fpca_fit = fpca_fit)
  pred_t395 <- out_samp_dyn_pred(df_new = df %>% filter(id==subjID & sind_inx<=395) %>%
                                  select(-eta_i),
                                fpca_fit = fpca_fit)
  pred_t595 <- out_samp_dyn_pred(df_new = df %>% filter(id==subjID & sind_inx<=595) %>%
                                  select(-eta_i),
                                fpca_fit = fpca_fit)

@@

```{r, laplace_ped}
# clean MLE out-of-sample prediction results
out_pred_clean_laplace <- function(subjID=1, mid=mid){
  # Prediction
  pred_t195 <- out_pred_laplace(fpca_fit = fpca_fit, df_new = df %>% filter(id==subjID & sind_inx<=195) %>% select(-eta_i))
  pred_t395 <- out_pred_laplace(fpca_fit = fpca_fit, df_new = df %>% filter(id==subjID & sind_inx<=395) %>% select(-eta_i))
  pred_t595 <- out_pred_laplace(fpca_fit = fpca_fit, df_new = df %>% filter(id==subjID & sind_inx<=595) %>% select(-eta_i))

```{r, results='hide'}
out_pred_lap_subj1 <- out_pred_clean_laplace(1, mid) %>% 
  left_join(df_true_eta %>% filter(id==1), by = "mid")
out_pred_lap_subj2 <- out_pred_clean_laplace(2, mid) %>% 
  left_join(df_true_eta %>% filter(id==2), by = "mid")
out_pred_lap_subj3 <- out_pred_clean_laplace(3, mid) %>% 
  left_join(df_true_eta %>% filter(id==3), by = "mid")
out_pred_lap_subj4 <- out_pred_clean_laplace(4, mid) %>% 
  left_join(df_true_eta %>% filter(id==4), by = "mid")
```

```{r, fig.width=12, fig.height=4}
# compare point prediction with in-sample
# subject 1
grid.arrange(
  out_pred_subj1 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 1 out-of-sample MLE"),

  out_pred_lap_subj1 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 1 out-of-sample Laplace"),

@@

# subject 2
grid.arrange(
  out_pred_subj2 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 2 out-of-sample MLE"),

  out_pred_lap_subj2 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 2 out-of-sample Laplace"),

@@

# subject 3
grid.arrange(
  out_pred_subj3 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 3 out-of-sample MLE"),

  out_pred_lap_subj3 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 3 out-of-sample Laplace"),

@@

# subject 4
grid.arrange(
  out_pred_subj4 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 4 out-of-sample MLE"),

  out_pred_lap_subj4 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 4 out-of-sample Laplace"),

@@

#### Laplace approximation

This method essential finds $\boldsymbol{\xi}$ that maximizes $p(\boldsymbol{Y}|\boldsymbol{\xi})$.

- <span style="color: red;"> Is it equivalent to posterior mean $E(\boldsymbol{\xi}|\boldsymbol{Y})$ under certain conditionss? (according to decision theory) </span>
- <span style="color: red;"> How to use Laplace Approximation to approximate integrals? </span>

https://search.r-project.org/CRAN/refmans/LaplacesDemon/html/LaplaceApproximation.html

@@

```{r, results='hide'}
# true scores:xi

# scores
in_score <- matrix(NA, nrow=N, ncol=4)
out_score_MLE <- matrix(NA, nrow=N, ncol=4)
out_score_laplace <- matrix(NA, nrow=N, ncol=4)

for(i in 1:N){
  in_score[i, ] <- in_samp_dyn_pred(fpca_fit=fpca_fit, 
                   df_part=df_pred_unique %>% filter(id==i & bin<=595))$score
  out_score_MLE[i,] <- out_samp_dyn_pred(df %>% filter(id==i & sind_inx<=595) %>%
                                  select(-eta_i),
                                  fpca_fit=fpca_fit)$score_out
  out_score_laplace[i,] <- out_pred_laplace(fpca_fit=fpca_fit, 
                                            df_new = df %>% filter(id==i & sind_inx<=595) %>% select(-eta_i))$score_out
}

# bias
in_bias <- in_score-xi
out_bias_MLE <- out_score_MLE-xi
out_bias_laplace <- out_score_laplace-xi
```

```{r, fig.height=4, fig.width=8}
grid.arrange(
  out_bias_MLE %>% data.frame() %>%
    pivot_longer(1:4, names_to = "PC", values_to = "Bias") %>%
    ggplot()+
    geom_boxplot(aes(x=PC, y=Bias))+
    stat_summary(aes(x=PC, y=Bias), fun = mean, geom = "point", col = "red")+
    labs(title = "Out-of-sample MLE"),

  out_bias_laplace %>% data.frame() %>%
    pivot_longer(1:4, names_to = "PC", values_to = "Bias") %>%
    ggplot()+
    geom_boxplot(aes(x=PC, y=Bias))+
    stat_summary(aes(x=PC, y=Bias), fun = mean, geom = "point", col = "red")+
    labs(title = "Out-of-sample laplace"),

@@

$$Var(\hat{\boldsymbol{\eta}}-\boldsymbol{\eta})=Var(\boldsymbol{\hat{f_0}-f_0})+\boldsymbol{\Phi\Gamma \Phi^T}+\hat{\sigma_{\epsilon}}^2$$

What happens to the variance of mean function? In practice it is estimated by smoothed population mean at each observation time.

Note:
